---
title: "Rexercise#3_fatingana"
author: "Kent A. Fatingana"
date: "2026-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Part A: Creating Vectors: c(), seq(), rep()  
```{r}
scores <- c(88, 92, 79, 95, 84)

subjects <- c("Math", "Science", "English", "History", "PE")
```


```{r}

odds <- seq(1, 19, by = 2)


numbers <- seq(0, 1, length.out = 5)
```

```{r}

rep("A", 4)


rep(c("A", "B"), each = 2)


rep(c("A", "B"), times = 2)
```


Question #1:

```{r}

min(age)
max(age)
age <- c(15, 28, 32, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22,
         37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 12, 42, 53,
         41, 51, 35, 24, 33, 8)
new_age <- c(age, 0, age)
```

a.How many data points?
Answer:There are 34 data points in the age vector.

b.Find the reciprocal of age?
Answer:The reciprocal of each value in age is calculated as 1 / age.

c.new_age contains 69 data points, with 0 in the middle.
Answer:new_age contains 69 data points, with 0 in the middle.

d: Find the minimum and the maximum value of age.
Answer:thw minimum age is 8 and the maximum is 57.

Question #2:

```{r}
rep(c("A", "B"), each = 2)
# "A" "A" "B" "B"

rep(c("A", "B"), times = 2)
# "A" "B" "A" "B"
```

rep(c("A","B"), each = 2) and rep(c("A","B"), times = 2) both produce a vector of length 4, but with different element orders.
Explain in your own words the difference between each and times.
Answer:The each is repeat each item while the times is repeat the whole sequence 

Question #3:
```{r}
seq(1, 20, by = 2)

seq(1, 20, length.out = 5)
```


Compare seq(1, 20, by = 2) with seq(1, 20, length.out = 5). What does each argument (by vs length.out) control, and when would you use one over the other in a real analysis?
Answer:The by argument controls the interval or step between each number, while the length.out is controls the total number of values in the sequence.


Question #4:
a
```{r}
vec_num <- seq(20, 60)
sum(vec_num)
```


b
```{r}
mean(vec_num)
```


c
```{r}
sum(vec_num)
```

d
```{r}
length(vec_num)
```

d: How many data points are there in vec_num?
Answer:41


part B;Subsetting (Accessing) Vectors

```{r}
grades <- c(88, 92, 79, 95, 84)
names(grades) <- c("Math", "Science", "English", "History", "PE")


grades[3]
# 79


grades[c(1, 4)]
# 88 95

grades[c("Science", "PE")]
# Science PE 
#      92  84
```


```{r}
grades[-c(1, 4)]
```

```{r}
grades[grades >= 85]
```

Question #5:

Negative indexing (grades[-1]) removes an element, while logical indexing (grades[grades >= 85]) filters by a condition.

Describe a real situation where you would need negative indexing, and a different situation where logical indexing would be the better tool.
Answer:Negative indexing: You could use it when you want to remove a specific item, such as removing a student’s grade from a list because the grade was entered incorrectly.
Logical indexing: You could use it when you want to select items based on a condition, such as finding all students who received a grade of 85 or higher.


```{r}
grades[10]
```
Question #6:
What happens if you run grades[10] (an index that doesn’t exist)? Run it, observe the result, and explain why R responds this way instead of throwing an error.
Answer:R returns NA result because grades only has 5 elements, so position 10 does not exist.

Question #7:
```{r}
Filter(function(i) { all(i %% c(3, 5, 7) != 0) }, seq(100))
```

Part C:Vector Operations and append()

```{r}
x <- c(1, 2, 3, 4)
y <- c(10, 20, 30, 40)

x + y
# [1] 11 22 33 44
x * 2
# [1] 2 4 6 8
```



```{r}
z <- c(1, 2, 3, 4, 5, 6)
short <- c(1, 0)

result <- z + short

print(result)
```


```{r}
fruits <- c("apple", "banana", "cherry")

fruits <- append(fruits, "date")
print(fruits)

# Insert "kiwi" between "apple" and "banana"
fruits <- append(fruits, "kiwi", after = 1)
print(fruits)
```
Question #8:
In the vectors-recycling chunk, z has 6 elements and short has only 2. R still produced an answer instead of an error.

Explain, in your own words, what R’s “recycling rule” did to short to make the addition work.
Answer:R’s recycling rule is repeated the short vector until it matched in the length of z 

```{r}
z <- c(1, 2, 3, 4, 5, 6)
z + c(1, 0, 1)
```


Question #9: Now try adding z (length 6) to c(1, 0, 1) (length 3) instead. Does it work cleanly?
Answer:yes its work cleanly because 6 is a multiple of 3 

Part D — Lists: Creating and Subsetting
```{r}
student <- list(
  name = "Maria Santos",
  age = 20,
  grades = c(88, 92, 79),
  is_enrolled = TRUE
)

print(student)
```


```{r}

grades <- student["grades"]


class(grades)
```

```{r}

grades <- student[["grades"]]


class(grades)
```

```{r}

student$name
```

```{r}

student[["grades"]][2]
```


Question #10:
```{r}
class(student["grades"])

class(student[["grades"]])
```
Question #10:

student["grades"] and student[["grades"]] look similar but return objects of different classes.
Run class() on each and explain the difference in what single-bracket [ ] versus double-bracket [[ ]] return from a list.
Answer:the single bracket[] is a lists of containing the selected element while the double bracket [[]] is a actual element of itself

Question #11:

Why do you think R makes a distinction between [ ] and [[ ]] for lists, but a vector only needs [ ]? Think about what a list can contain that a vector cannot.
Answer:The R makes this distinction because a list can contain different types of objects, such as numbers, text, vectors, or even other lists. The [ ] operator selects an element while keeping it as a list, while [[ ]] extracts the actual object inside the list.

Part E — Critical Thinking Challenge

```{r}
inventory <- list(
  item = c("Notebook", "Pen", "Eraser"),
  price = c(25.50, 10.00, 5.00),
  in_stock = c(TRUE, TRUE, FALSE)
)


print(inventory$item[inventory$in_stock == TRUE])
```


```{r}

inventory$item <- append(inventory$item, "Ruler")


inventory$price <- append(inventory$price, 15.00)


inventory$in_stock <- append(inventory$in_stock, TRUE)


print(inventory)
```


Question #12:

In challenge-1, you combined logical indexing on one vector (in_stock) to filter a different vector (item) inside the same list. Explain, step by step, why inventory$item[inventory$in_stock] works even though item and in_stock are stored as separate elements of the list.
Answer:inventory$item[inventory$in_stock] works because in_stock contains TRUE or FALSE values that match the positions of the items, so R keeps the items marked TRUE and removes those marked FALSE.

Question #13:

Suppose a classmate stores related data as three separate, unconnected vectors (item_names, item_prices, item_stock) instead of grouping them into one list like inventory. What risks or bugs could this create as the dataset grows or gets reordered? Argue for or against using a list (or, later, a data frame) instead of separate vectors.
Answer:I agree with using a list or, later, a data frame because keeping related data together reduces mistakes. With separate vectors, it is easy for the item names, prices, and stock values to become mismatched when data is added, removed, or reordered, which can cause incorrect information to be associated with an item. A list keeps the related vectors together, while a data frame is even better for larger datasets because each row represents one item and its related information.

Conclusion:

In 3–4 sentences, summarize the key differences between vectors and lists in R, and explain when you would choose one data structure over the other.
Answer:A vector in R stores elements of the same basic data type, such as numbers, characters, or logical values. A list can store different types of data, including vectors, numbers, text, and other lists, in one structure.I choose vector when i working with a simole collection of similaar values while I choose the list when i need to group the different types of values